Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-03-24 paper/paper.tex
thorough proofread 9 (into and outro) * reworded clunky paragraph on cplx-networks in the intro * rewritten the conclusion section

Inherently complex-valued nature of data in signal processing, specifically, acoustic or radio
frequency signal analysis, has been the key driver behind the adoption of $\cplx$-networks.
%
\citet{hirose_complex-valued_2009} argues that the key merit of $\cplx$-valued neural
networks is simultaneous phase rotation and amplitude adjustment, intrinsic to multiplication
in $\cplx$ field. Despite inherently higher arithmetic complexity, which stems from up to four
floating point operations per single $\cplx$ multiplication
$\cplx$-arithmetic reduces ineffective degrees of freedom in comparison with an analogous
double-dimensional $\real$-valued network. \citet{hirose_complex-valued_2009} lists numerous
applications related to acoustic and radio signal processing and demonstrates superiority of
$\cplx$-networks in the task of shallow landmine detection using ground penetrating radar
imaging. More recently they have been applied to music transcription and speech recognition
\citep{trabelsi_deep_2017}. \citet{arjovsky_unitary_2016} propose $\cplx$-valued unitary-RNNs
to deal with exploding~/~vanishing gradients and apply the recurrent models to long-term
dependence prediction tasks. \citet{wisdom_full-capacity_2016} expand on their work by optimizing
recurrent weights on the $\cplx^{n \times n}$ Stiefel manifold. And \citet{wolter_complex_2018}
investigate different $\cplx$-valued gating mechanisms for unitary-RNNs and conduct experiments
on human motion prediction and music transcription. \citet{yang_complex_2019} develop complex
transformer, with $\cplx$-attention and complex encoder-decoder, and apply it to music
transcription and wireless signal classification.

@@

We propose to use fully factorized complex Gaussian posterior approximation and $\cplx$
variant of the local reparameterization trick to extend Bayesian Dropout techniques to
$\cplx$-valued networks. In this section we describe the approximation and derive the
divergence penalties in \eqref{eq:elbo}.

@@

We study the compression-performance trade-off of the proposed Bayesian sparsification
method for $\cplx$-valued networks presented above on image classification and music
annotation tasks.
%
Following \citet{wolter_complex_2018} and \citet{trabelsi_deep_2017}, the class logit scores
are taken as real part of the complex-valued output of a network.
%
% image features
Since image data is not naturally $\cplx$-valued, we preprocess it by either embedding
$\real$-data into $\cplx$ assuming $\Im z = 0$ (\texttt{raw}), or applying the two-dimensional
Fourier Transform (\texttt{fft}), centering the lower frequencies. We do not train an
auxiliary model to synthesize the imaginary component from the real data, proposed by
\citet{trabelsi_deep_2017}.

@@

We conduct a moderately sized experiment on MNIST-like datasets of $28\times 28$ greyscale
images to study the compression-performance trade-off of the proposed $\cplx$-variational
dropout: MNIST \citep{lecun_gradient-based_1998}, KMNIST \citep{clanuwat_deep_2018},
EMNIST \citep{cohen_emnist_2017} and Fashion-MNIST \citep{xiao_fashion-mnist_2017}.
%
% dataset size
We deliberately use a {fixed} random subset of $10k$ images from the train split of each
dataset to fit the networks and measure the performance with classification \textit{accuracy}
score on the usual test split.
% Limiting the size was done primarily to test the regularization strength of compression
% and to facilitate quicker experiments.

@@

We conduct an experiment on the CIFAR10 classification dataset \citep{krizhevsky_learning_2009}
comprising $32\times 32$ color images of $10$ classes to study the compression-performance
trade-off of a moderately large deep network.
% same experimental setup as in MNIST, describe the VGG16 model and its cplx-valued modification
The experiment is set up similarly to MNIST-like experiment, except that we train $\real$
VGG16 network \citep{simonyan_very_2015} and its $\cplx$ variant, in which we have replaced
$\real$-valued layers with their $\cplx$-valued counterparts. $\real$ and $\cplx$ VGG16 have
the same feature sizes in intermediate layers.

@@

Despite having higher arithmetic complexity, complex-valued models provide built-in
amplitude-phase transformations and regularization \citep{hirose_complex-valued_2009}
and enjoy superior performance on naturally $\cplx$-valued datasets \citep{tarver_design_2019}.
In this study we address the issue of excessive arithmetic complexity of complex-valued
networks by extending the existing Bayesian compression techniques by \citet{kingma_variational_2015},
\citet{molchanov_variational_2017} and \citet{kharitonov_variational_2018} to $\cplx$-valued
parameters. To this end we have conducted a large numerical study of such networks to assess
the performance-compression rate trade-off and have achieved performance comparable to
the state-of-the-art in $\cplx$-valued networks on the MusicNet dataset.

Much like $\real$-valued variational dropout, the proposed $\cplx$-valued extension can be
straightforwardly enhanced to provide structured sparsity. The results of this study have
direct implications for real-time signal processing with embedded deep learning applications
both in terms of lower storage requirements and higher throughput stemming from fewer
floating-point multiplications.

>>>>

Inherently complex-valued nature of data in signal processing, specifically, acoustic or radio
frequency signal analysis, has been the key driver behind the adoption of $\cplx$-networks.
%
\citet{hirose_complex-valued_2009} argues that $\cplx$-valued networks merit from the
simultaneous phase rotation and amplitude adjustment, intrinsic to $\cplx$ arithmetic,
which reduce ineffective degrees of freedom in comparison with an double-dimensional
$\real$-valued networks. Their study lists numerous applications related to acoustic
and radio signal processing and demonstrates superiority of $\cplx$-networks in the
task of shallow landmine detection using ground penetrating radar imaging, in spite of
inherently higher arithmetic complexity of $\cplx$-networks, stemming from $3-4$ floating
point operations per single $\cplx$ multiplication. More recently $\cplx$-networks have
been applied to music transcription and speech recognition \citep{trabelsi_deep_2017}.
\citet{arjovsky_unitary_2016} propose $\cplx$-valued unitary-RNNs to deal with exploding~/~vanishing
gradients and apply the recurrent models to long-term dependence prediction tasks.
\citet{wisdom_full-capacity_2016} expand on their work by optimizing recurrent weights on
the $\cplx^{n \times n}$ Stiefel manifold. And \citet{wolter_complex_2018} investigate
different $\cplx$-valued gating mechanisms for unitary-RNNs and conduct experiments on
motion prediction and music transcription. \citet{yang_complex_2019} develop complex
transformer, with $\cplx$-attention and complex encoder-decoder, and apply it to music
transcription and wireless signal classification. \citet{tarver_design_2019} achieve
significantly lower out-of-band performance with a $\cplx$-valued network in Digital Signal
Predistortion task.

@@

We propose to use fully factorized complex Gaussian posterior approximation and $\cplx$
variant of the local reparameterization trick to extend Bayesian Dropout techniques to
$\cplx$-valued networks. In this section we describe the approximation and derive the
divergence penalties in \eqref{eq:elbo}. Similarly to $\real$-valued Variational Dropout,
the proposed $\cplx$-valued extension can be straightforwardly enhanced to provide
structured sparsity.

@@

We study the performance-compression trade-off of the proposed Bayesian sparsification
method for $\cplx$-valued networks presented above on image classification and music
annotation tasks.
%
Following \citet{wolter_complex_2018} and \citet{trabelsi_deep_2017}, the class logit scores
are taken as real part of the complex-valued output of a network.
%
% image features
Since image data is not naturally $\cplx$-valued, we preprocess it by either embedding
$\real$-data into $\cplx$ assuming $\Im z = 0$ (\texttt{raw}), or applying the two-dimensional
Fourier Transform (\texttt{fft}), centering the lower frequencies. We do not train an
auxiliary model to synthesize the imaginary component from the real data, proposed by
\citet{trabelsi_deep_2017}.

@@

We conduct a moderately sized experiment on MNIST-like datasets of $28\times 28$ greyscale
images to study the performance-compression trade-off of the proposed $\cplx$-variational
dropout: MNIST \citep{lecun_gradient-based_1998}, KMNIST \citep{clanuwat_deep_2018},
EMNIST \citep{cohen_emnist_2017} and Fashion-MNIST \citep{xiao_fashion-mnist_2017}.
%
% dataset size
We deliberately use a {fixed} random subset of $10k$ images from the train split of each
dataset to fit the networks and measure the performance with classification \textit{accuracy}
score on the usual test split.
% Limiting the size was done primarily to test the regularization strength of compression
% and to facilitate quicker experiments.

@@

We conduct an experiment on the CIFAR10 classification dataset \citep{krizhevsky_learning_2009}
comprising $32\times 32$ color images of $10$ classes to study the performance-compression
trade-off of a moderately large deep network.
% same experimental setup as in MNIST, describe the VGG16 model and its cplx-valued modification
The experiment is set up similarly to MNIST-like experiment, except that we train $\real$
VGG16 network \citep{simonyan_very_2015} and its $\cplx$ variant, in which we have replaced
$\real$-valued layers with their $\cplx$-valued counterparts. $\real$ and $\cplx$ VGG16 have
the same feature sizes in intermediate layers.

@@

In this study we have extended Variational Dropout methods of \citet{kingma_variational_2015},
\citet{molchanov_variational_2017} and \citet{kharitonov_variational_2018} to $\cplx$-valued
Bayesian networks. To validate the proposed technique we have conducted a large numerical
study of complex-valued networks with simple architectures in image recognition tasks to
assess the feasible performance-compression trade-off. At the cost of marginally lower performance,
we have achieved $50-100\times$ compression of the state-of-the-art deep $\cplx$-valued network
of \citet{trabelsi_deep_2017} on the MusicNet.

The results of this study have direct implications for real-time signal processing with
embedded deep learning applications both in terms of lower storage requirements and higher
throughput stemming from fewer floating-point multiplications, despite higher arithmetic
complexity of $\cplx$-valued networks.
